Optimal Symmetric Multimodal Templates and Concatenated Random Forests for Supervised Brain Tumor Segmentation (Simplified) with \textit{ANTsR}

### 2014-04-30 paper.tex
ENH:  More Brian comments.

%\paragraph{Stage 1 random forest} 
The stage 1 random forest training uses the asymmetry features, the seven posterior probability
tissue images output by the GMM
along with five additional geometric features based on the connected 
components of the segmentation image (distance image to tumor core,
physical volume image, volume to surface area ratio, eccentricity,
and elongation).
During the training phase (or RF model construction),
these feature images, $\left\{F_i: i \in GMM\right\}$, generated during Stage 1 are used to create
the Stage 1 RF model using the following relationship
\begin{align}
\label{eq:gmm}
 v_L \sim \sum_{i \in Asym} F_i + \sum_{i \in GMM} F_i + \sum_{i \in Misc} F_i
\end{align}
given in the standard \textit{R} notation of \cite{wilkinson1973}.  
In other words, $\bm{F}= [ \bm{F}_{Asym}, \bm{F}_{GMM},
\bm{F}_{Misc} ]$, where we column-wise concatenate the GMM, Asym and
Misc feature matrices.
The set of $\left\{F_i: i \in Asym\right\}$ are the feature images 
generated from the symmetric multivariate template 
described in the previous section and $\left\{F_i: i \in Misc\right\}$
are miscellaneous feature images described in Section \ref{sec:misc}.
Therefore, in total, the number of features is $k=84$ where there
are twelve GMM features for each of the four modalities ($n=48$), 
two asymmetry features for each modality ($n=8$), and six modality-specific 
miscellaneous geometric features ($n=24$) with an additional four feature
images.

@@

These spatial probability maps are passed as spatial priors to the
{\tt antsAtroposN4.sh} script in ANTs which couples the Atropos 
segmentation tool \citep{avants2011} with N4 bias correction \citep{tustison2010}.
The posterior probabilities that are output by Atropos and the
connected-component feature images for each modality (described in
section~\ref{sec:misc}) are then passed to the second stage random
forest model, along with the same asymmetry feature set based on our
symmetric templates.  
The second stage RF model is then produced similarly as the previous stage: 
\begin{align}
\label{eq:mapmrf}
\mathrm{Tissue}\,\,\mathrm{classes} \sim \sum_{i \in MAP\_MRF} F_i + \sum_{i \in Asym} F_i + \sum_{i \in Misc} F_i,
\end{align}
where $MAP-MRF$ represents the Stage 2 application of Atropos and the
other features are similar to those used in Stage 1.
The use of these MRF features improves the accuracy of the 
intensity-based modeling over the GMM approach by introducing
a spatial coherence constraint (the MRF prior) to those features 
in addition to 
providing a better initialization with the RF posteriors from 
the initial estimation stage (which lacked the MRF spatial prior).

>>>>

%\paragraph{Stage 1 random forest} 
The stage 1 random forest training uses the asymmetry features, the seven posterior probability
tissue images output by the GMM
along with five additional geometric features based on the connected 
components of the segmentation image (distance image to tumor core,
physical volume image, volume to surface area ratio, eccentricity,
and elongation).
During the training phase (or RF model construction),
these feature images, $\left\{F_i: i \in GMM\right\}$, generated during Stage 1 are used to create
the Stage 1 RF model using the following relationship
\begin{align}
\label{eq:gmm}
 v_L \sim \sum_{i \in Asym} F_i + \sum_{i \in GMM} F_i + \sum_{i \in Misc} F_i
\end{align}
given in the standard \textit{R} notation of \cite{wilkinson1973}.  
In other words, $\bm{F}= [ \bm{F}_{Asym}, \bm{F}_{GMM},
\bm{F}_{Misc} ]$, where we column-wise concatenate the GMM, Asym and
Misc feature matrices.
The set of $\left\{F_i: i \in Asym\right\}$ are the feature images 
generated from the symmetric multivariate template 
described in the previous section and $\left\{F_i: i \in Misc\right\}$
are miscellaneous feature images described in Section \ref{sec:misc}.
Therefore, in total, the number of features is $k=84$ where there
are twelve GMM features (7 posteriors plus 5 geometric features) for each 
of the four modalities ($n=48$), 
two asymmetry features for each modality ($n=8$), and six modality-specific 
miscellaneous geometric features ($n=24$) with an additional four feature
images.

@@

These spatial probability maps are passed as spatial priors to the
{\tt antsAtroposN4.sh} script in ANTs which couples the Atropos 
segmentation tool \citep{avants2011} with N4 bias correction \citep{tustison2010}.
The posterior probabilities that are output by Atropos and the
connected-component feature images for each modality (described in
section~\ref{sec:misc}) are then passed to the second stage random
forest model, along with the same asymmetry feature set based on our
symmetric templates.  
The second stage RF model is then produced similarly as the previous stage: 
\begin{align}
\label{eq:mapmrf}
v_L \sim \sum_{i \in MAP\_MRF} F_i + \sum_{i \in Asym} F_i + \sum_{i \in Misc} F_i,
\end{align}
where $MAP-MRF$ represents the Stage 2 application of Atropos and the
other features are similar to those used in Stage 1.
The use of these MRF features improves the accuracy of the 
intensity-based modeling over the GMM approach by introducing
a spatial coherence constraint (the MRF prior) to those features 
in addition to 
providing a better initialization with the RF posteriors from 
the initial estimation stage (which lacked the MRF spatial prior).  
In total, the number of features is for this final stage is also $k=84$ 
where the GMM features are replaced by their MAP-MRF counterparts
(7 posteriors plus 5 geometric features) for each of the four modalities ($n=48$).
The rest of the feature images from the first stage
(the two asymmetry features for each modality, six modality-specific 
miscellaneous geometric features, and four miscellaneous feature
images) are also included unaltered for this second stage.
